---
title: "MATH 4760 Exam 2"
author: "Liam Fruzyna"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---  
---

### 5) For the following four data sets, your objective is to come up with an appropriate ARIMA model (seasonal or non-seasonal).

- 1 Sulfer dioxide series, so2
- 2 Crude oil prices, oil
- 3 Global temperature data, gtemp
- 4 Johnson and Johnson earnings, jj

#### 1) Sulfer dioxide series, so2
##### a) Plot of the data
```{r}
library(astsa)
plot.ts(so2)
```

##### b) Box-Cox transformation if necessary, and the plot of the transformed data. Note that if a transformation is necessary, the transformed data must be used throughout.
Transformation is not necessary

##### c) Use appropriate techniques (if necessary) to remove trend and seasonal variations. Explain clearly what method(s) was used. Also submit the plot.

##### d) Plot of ACF and PACF. Explain clearly how you use them to determine a range of ARIMA model. Make sure to use differencing if necessary.
```{r}
acf2(so2)
```

Data set is not stationary so differencing is required.
```{r}
diff = diff(so2)
plot.ts(diff)
acf2(diff)
```

The ACF is used to determine q which would be 4 in this case and the PACF is used to determine p which would be 7 in this case.

##### e) Using certain criterion, determine an optimal ARMA(p, q) model.
```{r}
sarima(so2, 7, 1, 4)
```

##### f) Using hypothesis testing methods, check if certain parameters of the ARMA model can be removed. 

##### g) Performed diagnostic check for the model you obtained. Submit the appropriate plots. Make sure to use Box-Ljung statistics to test for white noise. If diagnostic check failed, adjust your model and start all over.  Compare all possible models you considered with AIC values and p-values of the Box-Ljung statistics. Determine the final model.
See e

##### h) Write the equation of the final model with clearly indicating the AR and MA coefficients. What is the estimate of the white noise variance? What does it tell you?
$ X_t+0.707X_{t-1}-0.081X_{t-2}-0.716X_{t-3}-0.089X_{t-4}-0.0937X_{t-5}-0.127X_{t-6}-0.008X_{t-7}=Z_t-0.178Z_{t-1}-0.635Z_{t-2}-0.767Z_{t-3}+0.580X_{t-4} $

##### i) Forecast the next 20 values, and submit the plot showing the data with forecast values together with their prediction intervals. State the forecasting values with their standard errors.
```{r}
sarima.for(so2, 20, 7, 1, 4)
```

#### 2) Crude oil prices, oil

##### a) Plot of the data
```{r}
plot.ts(oil)
```

##### b) Box-Cox transformation if necessary, and the plot of the transformed data. Note that if a transformation is necessary, the transformed data must be used throughout.
```{r}
log = log(oil)
plot.ts(log)
```

##### c) Use appropriate techniques (if necessary) to remove trend and seasonal variations. Explain clearly what method(s) was used. Also submit the plot.


##### d) Plot of ACF and PACF. Explain clearly how you use them to determine a range of ARIMA model. Make sure to use differencing if necessary.
```{r}
acf2(log)
```

Data set is not stationary so differencing is required.
```{r}
difflog = diff(log)
plot.ts(difflog)
acf2(difflog)
```

The ACF is used to determine q which would be 1 in this case and the PACF is used to determine p which would be 3 in this case.

##### e) Using certain criterion, determine an optimal ARMA(p, q) model.
```{r}
sarima(log, 3, 1, 1)
```

##### f) Using hypothesis testing methods, check if certain parameters of the ARMA model can be removed. 

##### g) Performed diagnostic check for the model you obtained. Submit the appropriate plots. Make sure to use Box-Ljung statistics to test for white noise. If diagnostic check failed, adjust your model and start all over.  Compare all possible models you considered with AIC values and p-values of the Box-Ljung statistics. Determine the final model.
See e

##### h) Write the equation of the final model with clearly indicating the AR and MA coefficients. What is the estimate of the white noise variance? What does it tell you?
$ X_t+0.322X_{t-1}+0.039X_{t-2}-0.118X_{t-3}=Z_t+0.495Z_{t-1} $

##### i) Forecast the next 20 values, and submit the plot showing the data with forecast values together with their prediction intervals. State the forecasting values with their standard errors.
```{r}
sarima.for(log, 20, 3, 1, 1)
```


---

#### 3) Global temperature data, gtemp

##### a) Plot of the data
```{r}
plot.ts(gtemp)
```

##### b) Box-Cox transformation if necessary, and the plot of the transformed data. Note that if a transformation is necessary, the transformed data must be used throughout.
Transformation was not necessary.

##### c) Use appropriate techniques (if necessary) to remove trend and seasonal variations. Explain clearly what method(s) was used. Also submit the plot.

##### d) Plot of ACF and PACF. Explain clearly how you use them to determine a range of ARIMA model. Make sure to use differencing if necessary.
```{r}
acf2(gtemp)
```

Data set is not stationary so differencing is required.
```{r}
diff = diff(gtemp)
plot.ts(diff)
acf2(diff)
```

The ACF is used to determine q which would be 1 in this case and the PACF is used to determine p which would be 3 in this case.

##### e) Using certain criterion, determine an optimal ARMA(p, q) model.
```{r}
sarima(gtemp, 3, 1, 1)
```

##### f) Using hypothesis testing methods, check if certain parameters of the ARMA model can be removed. 

##### g) Performed diagnostic check for the model you obtained. Submit the appropriate plots. Make sure to use Box-Ljung statistics to test for white noise. If diagnostic check failed, adjust your model and start all over.  Compare all possible models you considered with AIC values and p-values of the Box-Ljung statistics. Determine the final model.
See e

##### h) Write the equation of the final model with clearly indicating the AR and MA coefficients. What is the estimate of the white noise variance? What does it tell you?
$ X_t-0.070X_{t-1}+0.182X_{t-2}+0.130X_{t-3}=Z_t-0.581Z_{t-1} $

##### i) Forecast the next 20 values, and submit the plot showing the data with forecast values together with their prediction intervals. State the forecasting values with their standard errors.
```{r}
sarima.for(gtemp, 20, 3, 1, 1)
```


---

#### 4) Johnson and Johnson earnings, jj

##### a) Plot of the data
```{r}
plot.ts(jj)
```

##### b) Box-Cox transformation if necessary, and the plot of the transformed data. Note that if a transformation is necessary, the transformed data must be used throughout.
```{r}
log = log(jj)
plot.ts(log)
```

##### c) Use appropriate techniques (if necessary) to remove trend and seasonal variations. Explain clearly what method(s) was used. Also submit the plot.

##### d) Plot of ACF and PACF. Explain clearly how you use them to determine a range of ARIMA model. Make sure to use differencing if necessary.
```{r}
acf2(log)
```

Data set is not stationary so differencing is required.
```{r}
difflog = diff(log)
plot.ts(difflog)
acf2(difflog)
```

The ACF is used to determine q which would be 5 in this case and the PACF is used to determine p which would be 1 in this case.

##### e) Using certain criterion, determine an optimal ARMA(p, q) model.
```{r}
sarima(log, 1,1,5, 3,1,1, 4)
```

##### f) Using hypothesis testing methods, check if certain parameters of the ARMA model can be removed. 

##### g) Performed diagnostic check for the model you obtained. Submit the appropriate plots. Make sure to use Box-Ljung statistics to test for white noise. If diagnostic check failed, adjust your model and start all over.  Compare all possible models you considered with AIC values and p-values of the Box-Ljung statistics. Determine the final model.
See e

##### h) Write the equation of the final model with clearly indicating the AR and MA coefficients. What is the estimate of the white noise variance? What does it tell you?
$ X_t+0.243X_{t-1}=Z_t-0.360Z_{t-1}-0.061Z_{t-2}-0.303Z_{t-3}+0.831Z_{t-4}-0.334Z_{t-5} $

##### i) Forecast the next 20 values, and submit the plot showing the data with forecast values together with their prediction intervals. State the forecasting values with their standard errors.
```{r}
sarima.for(log, 20, 1,1,5, 3,1,1, 4)
```