---
title: "Final Part 1"
author: "Liam Fruzyna"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
# Liam Fruzyna
# MATH 4760
# Final Exam Part 1

library(astsa)
```

#### Consider the quarterly inflation data "qinfl"

#### 1) For ARIMA model forecast next the 12 quarters' inflations with graph sowing the forecast values with its 95% confidence intervals. 

```{r}
plot.ts(qinfl)
acf2(qinfl) # Needs to be differenced
dinfl = diff(qinfl)
acf2(dinfl)

sarima(qinfl, 1, 1, 3)
sarima.for(qinfl, 12, 1, 1, 3)

sarima(qinfl, 2, 1, 2, P=0, D=0, Q=2, S=4)
sarima.for(qinfl, 12, 2, 1, 2, P=0, D=0, Q=2, S=4)
```

#### 2) Fit the state-space model $x_t = \Phi x_{t-1} + z_t$ $y_t = x_t + v_t$ $i = 1,2,...,n$ where ${z_t}$ and ${v_t}$ are independent white noise with variances $\sigma ^2$ and $\sigma _v ^2$. assume that $x_0 ~ N(\mu _0, \Sigma _0)$
```{r}

```
##### a) Fit the smoothed $x_t ^n$ with figure showing the plot of msooth qinfl and its 95% confidence intervals.
```{r}
```
##### b) Forecast the next 12 quarters' inflations with a graph showing the forecast values with its 95% confidence intervals.
```{r}
```

### 3) Compare the results of the ARIMA and state-space models giving their pros and cons.
Any ARIMA model can be represented in a state-space form, however, only simple state-space models can be represented in ARIMA form.
ARIMA is good for approximations, it is never the exact model.
State-space requires writing down an actual model.
State-space allows for a greater variety of formulations, but models can get complicated.
State-space allows exact modelling but that can cause instability.