---
title: "Assignment 4"
author: "Liam Fruzyna"
output:
  pdf_document: 
    fig_height: 4
    keep_tex: yes
  html_document:
    df_print: paged
  html_notebook: default
---

```{r include=FALSE}
library(astsa)
abs = abs(nyse)
acf2(nyse)

library(fracdiff)
library(fGarch)
library(vars)
```

### 1) Consider the absolute values of the nyse returns data
#### a) Check using ACF whether abs(nyse) follows long memory or short memory
```{r}
acf2(abs)
```

Long memory, doesn't cut off or exponentially decay

#### b) Fit ARFIMA with appropriate order. Make sure to test the residuals.
```{r}
anyse = abs - mean(abs)
nyse.fd = fracdiff(anyse, nar=22, nma=6, M=30)
nyse.fd
fracdiff = diffseries(anyse, nyse.fd$d)
acf2(fracdiff)
Box.test(fracdiff, type='Ljung', lag=20)
```

### 2) Weekly crude oil spot prices in dollars per barrel are in oil
#### a Investigate whether the growth rate of the weekly oil prices exhibit GARCH behavior
```{r}
plot.ts(oil, ylab='Oil Prices')
acf2(oil)
```

#### b) Is the weekly growth rate white noise?
No, not white noise.

#### c) Fit an appropriate GARCH model
```{r}
fit = garchFit(~garch(1, 1), oil)

y = fit@sigma.t
plot(window(oil, start=2000, end=2010))
lines(window(oil-2*y, start=2000, end=2010), col=4)
lines(window(oil+2*y, start=2000, end=2010), col=4)
```

### 3) Let $S_t$ represent the monthly sales data, sales, and $L_t$ be a leading indicator lead
#### a) Plot the autocorrelation, acfs, or both sales and lead
```{r}
acf(sales)
acf(lead)
```

#### Also plot the cross-correlation ccf between sales and leads. Observer their behaviors.
```{r}
ccf(sales, lead)
```

#### b) Compute the difference of sales $\Delta S_t$ and difference of lead $\Delta L_t$.
```{r}
delta_s = diff(sales)
delta_l = diff(lead)
```

#### Repeat a for $\Delta S_t$ and $\Delta L_t$.
```{r}
acf(delta_s)
acf(delta_l)
ccf(delta_s, delta_l)
```

#### c) Fit the model $\Delta S_t = \beta _0 + \beta _1 \Delta L_{t-2} + \beta _2 \Delta L_{t-3} + x_t$, where ${x_t}$ is an ARMA process.
```{r}
fit = arima(delta_s, order=c(3,0,0), xreg=delta_l)
```

### 4) Consider the data set econ5. Conceptrate only on the unemployment, gnp, and consumption.
#### a) Fit an appropriate vector ARMA (VAR) to $x_t = (x_1t, x_2t, x_3t)$, where $x_1t = log(U_t)- \beta _0 - \beta _1 t$, $x_2t = G_t$, and $x_3t = C_t$. Meake sure to test for the residuals.
```{r}
plot.ts(econ5)
acf(econ5)

x=cbind(econ5$unemp, econ5$gnp, econ5$consum)
acf(x)
xnew = diff(x)
acf(xnew)

fit <- VAR(xnew, p=1)
summary(fit)
acf(resid(fit))
serial.test(fit, lags.pt=20)

VARselect(xnew, lag.max=10)

fit <- VAR(xnew, p=3)
summary(fit)
acf(resid(fit))
serial.test(fit, lags.pt=20)
```

#### b) Predict the unemployment of the next four quarters with 95% confidence intervals.
```{r}
fit.predict = predict(fit, n.ahead=4, ci=0.95)
fanchart(fit.predict)
```