---
title: "Math 4780 - Homework 2"
author: "Liam Fruzyna"
output:
  pdf_document: default
  html_notebook: default
---

### \#3.1 Consider the NFL data in table B.1
Load in dataset
```{r}
library(MPV)
nfl = table.b1
```

#### a) Fit a multiple linear regression model relating y to x2, x7, and x8
```{r}
model <- lm(y ~ x2 + x7 + x8, data=nfl)
summary <- summary(model)
model$coefficients
```

$\hat{y} = -1.8 + 0.0036x_2 + 0.196x_7 - 0.0048x_8$

#### b) Construct the anova table and test for significance
```{r}
anova(model)
```
All 3 variables are significant, their p-values are very small.

#### c) Calculate t statistics for testing the hypotheses
```{r}
summary
```

~~~
Coefficient  T-Val P-Value
x2           5.177 2.66e-05
x7           2.198 0.037815
x8          -3.771 0.000938
~~~

#### d) Calculate R^2 and adjusted R^2 values
```{r}
summary$r.squared
summary$adj.r.squared
```

$R^2 = 79\%$ and Adjusted $R^2 = 76\%$

#### e) Using the partial F test determine the contribution of x7 to the model
```{r}
reduced <- lm(y ~ x2 + x8, data=nfl)
anova(reduced, model)
```

F-Value is 4.83, it is significant at 95% bounds. This test statistic is the square of the t-statistic.

### \#3.3 Using 3.1
#### a) 95% CI on B7
```{r}
confint(model, 'x7', level=0.95)
```

The 95% confidence interval of $\beta_7$ is 0.012 to 0.376

#### b) 95% CI on mean number of games won when x2=2300 x7=56 and x8=2100
```{r}
predict(model, data.frame(x2=2300, x7=56, x8=2100), interval='confidence')
```

The 95% confidence interval of y is 6.436 to 7.997

### \#3.9 Consider the data in table B.6
Load in dataset
```{r}
library(MPV)
NbOCl <- table.b6
```

#### a) Fit a multiple linear regression model relating y to x1 and x4
```{r}
model <- lm(y ~ x1 + x4, data=NbOCl)
summary <- summary(model)
model$coefficients
```

$\hat{y} = 0.005 - 0.345x_1 -0.0001x_4$

#### b) Test for significance of regression
```{r}
anova <- anova(model)
anova
mean(anova$`F value`[1:2])
```

The F value is 24.66 which is significant to 95%

#### c) Calculate R^2 and adjusted R^2
```{r}
summary$r.squared
summary$adj.r.squared
```

$R^2 = 66\%$ and Adjusted $R^2 = 64\%$

#### d) Determine the contribution of x1 and x4 to the model with t-tests. Are they both necessary?
```{r}
summary
```

$x_1$ is definitely significant at its p-value is very low (significant to 99.9%), however, $x_4$ is not significant as its p-value is very high.

#### e) Is multicollinearity a concern?
```{r}
library(car)
vif(model)
```
No multicollinearity is not a concern because VIF is low for both variables.

### \#9 Prove H and I-H are idempotent
$HH = X(X^TX)^{-1}X^TX(X^TX)^{-1}X^T$

$HH = X(X^TX)^{-1}X^T$

$HH = H$

$(I - H)(I - H) = I - H - H + HH$

$(I - H)(I - H) = I - H - H + H$

$(I - H)(I - H) = (I - H)$