---
title: "Math 4780 - Homework 4"
author: "Liam Fruzyna"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

### \#5.2
Dataset setup
```{r}
temps <- seq(273, 373, 10)
pressures <- c(4.6, 9.2, 17.5, 31.8, 55.3, 92.5, 149.4, 233.7, 355.1, 525.8, 760)
vp <- data.frame(Temperature=temps, Vapor_Pressure=pressures)
```

#### a. Plot a scatter diagram. Does it seem likely that a straight-line model will be adequate?
```{r}
plot(vp, main='Vapor Pressure vs Temperature')
```

No, it does not appear that a straight line mode will be adequate because the data is non-linear.

#### b. Fit the straight-line model. Compute the summary statistics and the residual plots. What are your conclusions regarding the model adequacy?
```{r}
lm_vp <- lm(Vapor_Pressure ~ Temperature, data=vp)
summary(lm_vp)
plot(lm_vp, which=1)
```

The model is not accurate.

#### c. From physical chemistry the Clausius-Clapeyron equation states that $ln(p_v) \propto -\frac{1}{T}$. Repeat part b using the appropriate transformation based on this information.
```{r}
trans_vp <- data.frame(Temperature=log(temps), Vapor_Pressure=log(pressures))
plot(trans_vp, main='Vapor Pressure vs Temperature')

lm_trans_vp <- lm(Temperature ~ Vapor_Pressure, data=trans_vp)
summary(lm_trans_vp)
plot(lm_trans_vp, which=1)
```

There is some improvement in the model with a log transformation.

### \#5.5
Dataset setup
```{r}
defects <- c(13, 16.1, 14.5, 17.8, 22, 27.4, 16.8, 34.2, 65.6, 49.2, 66.2, 81.2, 87.4, 114.5)
weeks <- 4:17
glass <- data.frame(Defects_Per_10k=defects, Weeks=weeks)
plot(glass, main='Defects per 10K Units by Week')
```

#### a. Fit a straight-line regression model to the data and perform the standard tests for model adequacy.
```{r}
lm_glass <- lm(Defects_Per_10k ~ Weeks, data=glass)
summary(lm_glass)
plot(lm_glass, which=c(1,2))
```

There is a slight non-linear pattern to the residuals.

#### b. Suggest an appropriate transformation to eliminate the problems encountered in part a. Fit the transformed model and check for adequacy.
```{r}
trans_glass <- data.frame(Defects_Per_10k=log(defects), Weeks=weeks)
plot(trans_glass, main='Defects per 10K Units by Week')

lm_trans_glass <- lm(Defects_Per_10k ~ Weeks, data=trans_glass)
summary(lm_trans_glass)
plot(lm_trans_glass, which=c(1,2))
```

The model appears to have improved by adding the natural log.

### \#6.2 Perform a thorough influence analysis of the property valuation data given in B.4. Discuss your results.
```{r}
library(MPV)
prop <- table.b4
lm_prop <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8, data=prop)
cooks.distance(lm_prop)
```

No points appear to be influential.

### \#6.10 Formally show that $D_i=\frac{r_i}{p}\frac{h_{ii}}{1-h_{ii}}$
Using $\hat\beta - \hat\beta_{(i)} =  \frac{(X'X)^{-1} x_i e_i}{1-h_{ii}}$

$D_i = \frac{(\hat\beta_{(i)} - \hat\beta)' X'X (\hat\beta_{(i)} - \hat\beta)}{pMS_{Res}}$
$= \frac{x_i(X'X)^{-1}X'X(X'X)^{-1}x_ie_i^2}{(1-h_{ii})^2pMS_{Res}}$
$= (\frac{e_i}{1-h_{ii}})^2 (\frac{h_{ii}}{pMS_{Res}})$
$= \frac{e_i^2}{MS_{Res}(1-h_{ii})} \frac{1}{p} \frac{h_{ii}}{1-h_{ii}}$
$= \frac{r_i^2}{p} \frac{h_{ii}}{1-h_{ii}}$

### \#6.15 Table B.14 contains data converning the transient points of an electronic inverter. Fit a regression model to all 25 observations but only use $x_1 - x_4$ as the regressors. Investigate this model for influential observations and comment on your findings.
```{r}
invt <- table.b14
lm_invt <- lm(y ~ x1 + x2 + x3 + x4, data=invt)
cooks.distance(lm_invt)
```

Points 2 and 4 are most influential.